Weighted Visibly Pushdown Automata\\ Application to Automated Music Transcription

### 2021-05-01 main.tex
1-best algo update

\def\SWT{\textsf{SWT}\xspace}
\def\SWA{\textsf{SWA}\xspace}
\def\SWVPA{\textsf{SWVPA}\xspace}
\def\weight{\mathsf{weight}}
\def\wei{\mathsf{w}}
\def\init{\mathsf{in}}
\def\final{\mathsf{out}}
\newcommand{\call}[1]{\ensuremath #1} %{\ensuremath \langle_{#1}}
\newcommand{\return}[1]{\ensuremath #1} %{\ensuremath {}_{#1}{\rangle}} % $\prescript{}{a}{)}$
\def\Sigmai{{\Sigma_\mathsf{i}}}
\def\Sigmac{{\Sigma_\mathsf{c}}}
\def\Sigmar{{\Sigma_\mathsf{r}}}
\def\Deltai{{\Delta_\mathsf{i}}}
\def\Deltac{{\Delta_\mathsf{c}}}
\def\Deltar{{\Delta_\mathsf{r}}}
\def\Phii{{\Phi_\mathsf{i}}}
\def\Phic{{\Phi_\mathsf{c}}}
\def\Phir{{\Phi_\mathsf{r}}}
\def\Phicr{{\Phi_\mathsf{cr}}}
\def\weii{{\wei_\mathsf{i}}}
\def\weic{{\wei_\mathsf{c}}}
\def\weir{{\wei_\mathsf{r}}}
\def\weie{{\wei_\mathsf{e}}}
\newcommand{\config}[2]{\ensuremath \genfrac{[}{]}{0pt}{}{#1}{#2}} 
\newcommand{\opluseq}{\ensuremath\mathrel{\oplus}=}

@@

\title{Weighted Visibly Pushdown Automata and Automated Music Transcription}
\author{Florent Jacquemard}
%\institute{INRIA \& CNAM, Paris, France\\
%\email{florent.jacquemard@inria.fr}}

@@

weighted parsing

principle of quantitative parsing / parsing of words over infinite alphabet:

@@

\noindent
\quad\textbf{extract} from $P$ $\< q_1, \sigma, q_2>$ with a minimal $d(q_1, \sigma, q_2)$ wrt $\leq_\oplus$.

@@

\noindent
\qquad
\(\begin{array}{rrcl}
 & d(q_1, \sigma, q_3) & \opluseq &
   d(q_1, \sigma, q_2) \otimes \displaystyle\bigoplus_{a \in \Sigmai} \weii(q_2, a, q_3)\\
%
%  & d(q_0, \sigma, q_2) & \opluseq & 
%    \displaystyle\bigoplus_{a \in \Sigmai} \weii(q_0, a, q_1) \otimes d(q_1, \sigma, q_2)\\
%
\mathrm{if~} \sigma = \top & d(q_0, \top, q_3) & \opluseq &
d(q_1, \sigma, q_2) \otimes
%\displaystyle\bigoplus_{p \in P} 
\displaystyle\bigoplus_{{\call{c}} \in \Sigmac} 
\displaystyle\bigoplus_{{\return{r}} \in \Sigmar} \phi({\call{c}}, {\return{r}})\\
\mathrm{and} &  d(q_0, \bot, q_3) & \opluseq &
d(q_1, \sigma, q_2) \otimes
%\displaystyle\bigoplus_{p \in P} 
\displaystyle\bigoplus_{{\call{c}} \in \Sigmac} 
\displaystyle\bigoplus_{{\return{r}} \in \Sigmar} \phi({\call{c}}, {\return{r}})\\
& & & \mathrm{where~} \phi = \weic(q_0, q_1, p) \otimes \weir(q_2, p, q_3)\\
%
\mathrm{if~} \sigma = \bot & d(q_1, \bot, q_3) & \opluseq &
 d(q_1, \sigma, q_2) \otimes \displaystyle\bigoplus_{\return{r} \in \Sigmar} \weie(q_2, {\return{r}}, q_3)\\
%
%  & d(q_1, \sigma, q_3) & \opluseq &
%  d(q_1, \sigma, q_2) \otimes d(q_2, \sigma, q_3)
\end{array}
\) 
\end{algo}

@@

\subsection{Representation of Music Performances}
A musical performance is represented symbolically as a finite sequence of timestamped events.
similar to piano roll representation~\cite{Muller15fundamentals} chap.1

\noindent
We consider an infinite alphabet $\Sigma$ of MIDI-like 
events made of:
\begin{itemize}
\item a timestamp in RTU\\ %(real or int = nb samples ?)
      or IOI in RTU
\item a pitch value in 0..127
\item \textsf{on} | \textsf{off} flag
\item a velocity value in 0..127
\end{itemize}

\subsection{Representation of Music Scores}
we consider here the case of monodies for the sake of simplificity.

A music score is represented as a structured word
= sequence of quantified events + markups,
see nested words~\cite{AlurMadhusudan09nested}.

We consider an alphabet $\Delta$, every symbol of which is 
composed of a tag, in a finite set $\Xi$, 
and an IOI duration value.
Moreover, $\Delta$ is partitioned into 
$\Delta = \Deltai \uplus \Deltac \uplus \Deltar$, 
like in Section~\ref{section:SWVPA}.

\noindent
The elements of $\Deltai$ represent events:
% (infinite alphabet of internal symbols) made of:
whose tags are one of:
\begin{itemize}
\item continuation (0) : tie or dot
\item note, grace note (pitch) or chord (pitch+)
\item rest
\item ...
\end{itemize}

@@

>>>>

\def\SWT{\textsf{SWT}\xspace}
\def\SWA{\textsf{SWA}\xspace}
\def\SWVPA{\textsf{SWVPA}\xspace}
\def\weight{\mathsf{weight}}
\def\wei{\mathsf{w}}
\def\init{\mathsf{in}}
\def\final{\mathsf{out}}
\newcommand{\call}[1]{\ensuremath #1} %{\ensuremath \langle_{#1}}
\newcommand{\return}[1]{\ensuremath #1} %{\ensuremath {}_{#1}{\rangle}} % $\prescript{}{a}{)}$
\def\Sigmai{{\Sigma_\mathsf{i}}}
\def\Sigmac{{\Sigma_\mathsf{c}}}
\def\Sigmar{{\Sigma_\mathsf{r}}}
\def\Deltai{{\Delta_\mathsf{i}}}
\def\Deltac{{\Delta_\mathsf{c}}}
\def\Deltar{{\Delta_\mathsf{r}}}
\def\Phii{{\Phi_\mathsf{i}}}
\def\Phic{{\Phi_\mathsf{c}}}
\def\Phir{{\Phi_\mathsf{r}}}
\def\Phicr{{\Phi_\mathsf{cr}}}
\def\weii{{\wei_\mathsf{i}}}
\def\weic{{\wei_\mathsf{c}}}
\def\weir{{\wei_\mathsf{r}}}
\def\weie{{\wei_\mathsf{e}}}
\newcommand{\config}[2]{\ensuremath \genfrac{[}{]}{0pt}{}{#1}{#2}} 
\newcommand{\opluseq}{\ensuremath\mathrel{\oplus}=}
\newcommand{\ioi}[1]{\mathsf{ioi}({#1})}

@@

% Parsing over infinite alphabet as optimal alignment computation 
% as edit-distance between string and language
% Symbolic Parsing
\title{Weighted Visibly Pushdown Automata\\ Application to Automated Music Transcription}
\author{Florent Jacquemard}
%\institute{INRIA \& CNAM, Paris, France\\
%\email{florent.jacquemard@inria.fr}}

@@

problem
quantitative parsing 
or symbolic parsing
parsing of words over infinite input alphabet.

generalizes weighted parsing: 
finding the best derivation for a weighted grammar.

model of weighted CFG over infinite set of terminal symbol
(but finite set non-terminals and production rules).
in production rules, terminal symbols appear as variables (parameters)
and the weight associated to the production rule is a function of these variables.
It is the approach of Symbolic Automata, except that the domain of weight values 
is not restricted to be Boolean, like  for the guards in the rules of SA, 
but can be an arbitrary commutative semiring (assuming some restrictions).

wrt an edit distance

we consider a strictly restricted model of weighted grammars : VPA 
and general edit-distances that can be defined by a weighted word transducer

@@

\noindent
\quad\textbf{extract} $\< q_1, \sigma, q_2>$ from $P$ 
 such that $d(q_1, \sigma, q_2)$ is minimal wrt $\leq_\oplus$.

@@

\noindent
\qquad
\(\begin{array}{rrcl}
 & d(q_1, \sigma, q_3) & \opluseq &
   d(q_1, \sigma, q_2) \otimes \displaystyle\bigoplus_{a \in \Sigmai} \weii(q_2, a, q_3)\\
%
%  & d(q_0, \sigma, q_2) & \opluseq & 
%    \displaystyle\bigoplus_{a \in \Sigmai} \weii(q_0, a, q_1) \otimes d(q_1, \sigma, q_2)\\
%
\mathrm{if~} \sigma = \top & d(q_0, \top, q_3) & \opluseq &
d(q_1, \sigma, q_2) \otimes
%\displaystyle\bigoplus_{p \in P} 
\displaystyle\bigoplus_{{\call{c}} \in \Sigmac} 
\displaystyle\bigoplus_{{\return{r}} \in \Sigmar} \eta({\call{c}}, {\return{r}})\\
\mathrm{and} &  d(q_0, \bot, q_3) & \opluseq &
d(q_1, \sigma, q_2) \otimes
%\displaystyle\bigoplus_{p \in P} 
\displaystyle\bigoplus_{{\call{c}} \in \Sigmac} 
\displaystyle\bigoplus_{{\return{r}} \in \Sigmar} \eta({\call{c}}, {\return{r}})\\
& & & \mathrm{where~} \eta = \weic(q_0, q_1, p) \otimes \weir(q_2, p, q_3)\\
%
\mathrm{if~} \sigma = \bot & d(q_1, \bot, q_3) & \opluseq &
 d(q_1, \sigma, q_2) \otimes \displaystyle\bigoplus_{\return{r} \in \Sigmar} \weie(q_2, {\return{r}}, q_3)\\
%
& d(q_0, \bot, q_2) & \opluseq & 
d(q_0, \bot, q_1) \otimes d(q_1, \sigma, q_2), \mathrm{if~} \< q_0, \bot, q_1> \notin P \\
%
\mathrm{if~} \sigma = \top & d(q_0, \top, q_2) & \opluseq & 
d(q_0, \top, q_1) \otimes d(q_1, \sigma, q_2), \mathrm{if~} \< q_0, \top, q_1> \notin P \\
%
& d(q_1, \bot, q_3) & \opluseq & 
   d(q_1, \sigma, q_2) \otimes d(q_2, \bot, q_3), \mathrm{if~} \< q_2, \bot, q_3> \notin P \\
%
\mathrm{if~} \sigma = \top & d(q_1, \top, q_3) & \opluseq & 
    d(q_1, \sigma, q_2) \otimes d(q_2, \top, q_3), \mathrm{if~} \< q_2, \top, q_3> \notin P \\
\end{array}
\) 
\end{algo}

@@

\subsection{Representation of Music Performances}
We consider symbolic representation of musical performances, as finite sequences of events.
It corresponds to the concrete case of a MIDI file~\cite{SMF} 
recorded  from an electronic keyboard, 
or the output of a transcription from audio files~\cite{Benetos18AMTsurvey}.
%
For the sake of simplicity, we shall only consider here the case of monophonic performances, 
where at most one note is sounding at a time -- the approach however extends to the polyphonic case.

A music performance is a finite sequence of events in a set $\Sigma$.
Every event $e \in \Sigma$ has attributes such from a finite domain, 
like a number of key for a note 
or a flag indicating that it is a rest 
(\textsf{ON} and \textsf{OFF} messages in~\cite{SMF})
and a velocity value (0..127 in~\cite{SMF})).
%This representation is similar to the piano roll ~\cite{Muller15fundamentals} chap.1. 
Moreover, it contains a RTU value $\ioi{e}$ (real number) 
representing the time distance to the next event, or the to the end of performance 
(also called \emph{inter-onset interval}).

\subsection{Representation of Music Scores}
Music score are represented as structured words
made of quantified events and parenthesized markups,
akin of nested words~\cite{AlurMadhusudan09nested}.

We consider an alphabet $\Delta$, every symbol of which is 
composed of a tag, in a finite set $\Xi$, 
and an MTU (rational) IOI duration value.
The alphabet $\Delta$ is partitioned into 
$\Delta = \Deltai \uplus \Deltac \uplus \Deltar$, 
like in Section~\ref{section:SWVPA}.
%
\noindent
The symbols of $\Deltai$ represent events:
% (infinite alphabet of internal symbols) made of:
with tags indicate a new note or grace-note (with null IOI), 
a rest or the continuation of the previous note (tie or dot).
The elements of $\Deltac \uplus \Deltar$ are matched
and are the markups for describing the structure of the score.

@@

tempo value in a finite domain (e.g. 30..300 bpm)
can be fixed 
or recomputed by the transducer when reading each event, 
according to a perceptive/cognitive model of tempo 
such as~\cite{LargeJones99tempo}
(also used in the context of score following~\cite{Cont10TPAMI}).
% we wont detail here.
